Chapter 17 Lab - Random Forests and Boosting

### 2022-10-17 labs/Chapter 17 Lab.Rmd
Finish lab

```{r}

```

>>>>

```{r}
set.seed(495)
n <- nrow(Hitters)
train_index <- sample(1:n, 0.5 * n)
test_index <- setdiff(1:n, train_index)

hitters_train <- Hitters[train_index, ] 
hitters_test <- Hitters[test_index, ]

```
```{r}
tree_hitters <- tree(Salary ~ ., hitters_train)

summary(tree_hitters)

single_tree_estimate <- predict(tree_hitters, 
                               newdata = hitters_test)
ggplot() + 
    geom_point(aes(x = hitters_test$Salary, y = single_tree_estimate)) +
    geom_abline()

mean((single_tree_estimate - hitters_test$Salary)^2)
```
The MSE for a single tree is 142,426.9.

```{r}
# bagging

set.seed(495)
bag_hitters <- randomForest(Salary ~ ., 
                          data = hitters_train, 
                          mtry = 13, 
                          importance = TRUE)
bag_hitters

bagged_estimate <- predict(bag_hitters, 
                          newdata = hitters_test)

ggplot() + 
    geom_point(aes(x = hitters_test$Salary, y = bagged_estimate)) +
    geom_abline()

mean((bagged_estimate - hitters_test$Salary)^2)

```
Once we fit a model with bagging, the MSE decreased to 102,961.4 which is much lower than the single tree.

```{r}
# random forest

set.seed(495)
rf_hitters <- randomForest(Salary ~ ., 
                         data = hitters_train, 
                         mtry = 6, 
                         importance = TRUE)

random_forest_estimate <- predict(rf_hitters, 
                                 newdata = hitters_test)

mean((random_forest_estimate - hitters_test$Salary)^2)

varImpPlot(rf_hitters)
```
The MSE of the random forest model with `mtry` = 6 is 105,432.7 which is a bit higher than the bagging model.

```{r}
# boosting

set.seed(495)
boost_hitters <- gbm(Salary ~ ., 
                   data = hitters_train, 
                   distribution = "gaussian", 
                   n.trees = 5000, 
                   interaction.depth = 4)

summary(boost_hitters)

boost_estimate <- predict(boost_hitters, 
                         newdata = hitters_test, 
                         n.trees = 5000)

mean((boost_estimate - hitters_test$Salary)^2)
```
```{r}
set.seed(495)
boost_hitters2 <- gbm(Salary ~ ., 
                   data = hitters_train, 
                   distribution = "gaussian", 
                   n.trees = 5000, 
                   interaction.depth = 4,
                   shrinkage = 0.01, 
                   verbose = F)

summary(boost_hitters2)

boost_estimate2 <- predict(boost_hitters2, 
                         newdata = hitters_test, 
                         n.trees = 5000)

mean((boost_estimate2 - hitters_test$Salary)^2)

```
The boosting model has an MSE of 107,212.3 which is a bit higher than the other models
